---
format:
  pdf:
    toc: false
    number-sections: false
    geometry: "margin=0.5in"
    fontsize: 11pt
execute:
  echo: false
  message: false
  warning: false
knitr:
  opts_chunk:
    fig.align: "center"
---

**INVESTIGATING THE INFLUENCE OF HOST AGE ON MURINE MICROBIAL RESPONSES TO DIETARY CONTAMINANTS**

Amon Cox, Evelyn Callaway, and Arul Jayaraman\*

Artie McFerrin Department of Chemical Engineering, Texas A&M University, College Station, Texas, United States of America, 77843

\*Correspondence: arulj@tamu.edu

# Supplementary Material

## Table S1 Dunn’s Test for 100 &micro;M BPS Age 22-Weeks Alpha Diversity Scores

Bonferroni-corrected *p* values for each exposure and culture day by alpha-diversity index.

```{r}

library(readr)
library(knitr)
library(dplyr)
library(stringr)

bps_high_wk22_DunnTest <- readr::read_csv(file.path("output", "tables", "8_16s_alphaDiv_DunnTest_BPS_high_wk22.csv")) |>
    mutate(index = str_replace_all(index, c("1" = "Shannon", "2" = "Inverse Simpson"))) |>
    rename(`Alpha-Diversity Index` = index, `Adjusted p value` = P.adj) |>
    select(-Z, -P.unadj)

knitr::kable(bps_high_wk22_DunnTest, digits = 3)
```

{{< pagebreak >}}

## Table S2 Dunn’s Test for 100 &micro;M DINP Age 22-Weeks Alpha Diversity Scores

Bonferroni-corrected *p* values for each exposure and culture day by alpha-diversity index.

```{r}

dinp_high_wk22_DunnTest <- readr::read_csv(file.path("output", "tables", "8_16s_alphaDiv_DunnTest_DINP_high_wk22.csv")) |>
    mutate(index = str_replace_all(index, c("1" = "Shannon", "2" = "Inverse Simpson"))) |>
    rename(`Alpha-Diversity Index` = index, `Adjusted p value` = P.adj) |>
    select(-Z, -P.unadj)

knitr::kable(dinp_high_wk22_DunnTest, digits = 3)
```

{{< pagebreak >}}

## Table S3 Dunn’s Test for 100 &micro;M DINP Age 40-Weeks Alpha Diversity Scores

Bonferroni-corrected *p* values for each exposure and culture day by alpha-diversity index.

```{r}

dinp_high_wk40_DunnTest <- readr::read_csv(file.path("output", "tables", "8_16s_alphaDiv_DunnTest_DINP_high_wk40.csv")) |>
    mutate(index = str_replace_all(index, c("1" = "Shannon", "2" = "Inverse Simpson"))) |>
    rename(`Alpha-Diversity Index` = index, `Adjusted p value` = P.adj) |>
    select(-Z, -P.unadj)

knitr::kable(dinp_high_wk40_DunnTest, digits = 3)
```

{{< pagebreak >}}

## Table S4 Pairwise Comparisons for 100 &micro;M Exposure Beta-Diversity Scores

Adjusted *p* values for multilevel pairwise comparisons of Bray-Curtis dissimilarities for 16S microbial abundances. Each comparison is between a 100 &micro;M exposure and culture day across the host ages.

```{r}

pairwise_ls <- list.files(path = file.path("output", "tables"),
    pattern = "^8_16s_betaDiv_pairwise_",
    full.names = TRUE
) |>
    purrr::set_names(~ basename(.x) %>% sub("^8_16s_betaDiv_pairwise_(.*)\\.csv$", "\\1", .)) |> 
    purrr::map(.f = read_csv)

pairwise_df <- pairwise_ls |>
    bind_rows(.id = "dataset") |>
    select(-p_value) |>
    mutate(
        dataset = gsub("_high_", paste0(" 100", expression("\U00B5"), "M "), dataset),
        group1 = gsub("bps|dinp", "exposure", group1),
        group2 = gsub("bps|dinp", "exposure", group2)
    ) |>
    select(-R2) |>
    tidyr::pivot_wider(
        names_from = dataset,
        values_from = p_adj
    )

knitr::kable(pairwise_df, digits = 3)
```

{{< pagebreak >}}

## Table S5 Age-Dependent, Exposure-Differentiating Pathways

Putative metabolic pathways from LC-MS profiles of microbial cultures (n = 7) that were significantly differentially abundant (*q* < .05, absolute log~2~ fold change > 1) by exposure type and across host age (FDR < .05). Each row represents a specific pathway that met these criteria at a given culture day within a dataset (exposure type, dose level, and host age at time of culture inoculation). ^1^The fold change of the log~2~ pathway intensities (arbitrary units) between exposed and control cultures. ^2^A positive slope indicates the pathway increased in baseline intensity (abundance) within the inocula samples as the mice aged, whereas a negative slope indicates the baseline abundance decreased with host age. ^3^The number of LC-MS peaks putatively identified as chemical compounds which belong to the given pathway in the given dataset for that comparison.

\tiny
```{r}

age_exp_paths <- read_tsv(file.path("output", "tables", "13_age_exp_pathways.tsv")) |>
    dplyr::filter(abs(log2FC_exp) > 1) |>
    select(dataset, culture_day, pathwayID, log2FC_exp, slope, rep_features, pathway) |>
    arrange(dataset, pathwayID, culture_day) |>
    rename(`culture day` = culture_day, "\u00B9log\u2082FC" = log2FC_exp, "\u00B2age trend" = slope, "\u00B3features" = rep_features)

knitr::kable(age_exp_paths)
```
\normalsize
